quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-09-18 paper-quollr.Rmd
Updated vertical spacing

The coordinate limits of the embedding (`lim1` and `lim2`) are used to compute the aspect ratio between the two axes, which informs vertical spacing. The function then calls `calc_bins_y()`, a helper function that determines the appropriate number of hexagonal rows (`bin2`) and the width of each hexagon (`a1`) given the specified number of bins along the x-axis (`b1`) and buffer (`q`).

Then, the centroids are computed iteratively. The x-coordinates for centroids in odd-numbered rows are initialized as a sequence spaced by the hexagon width. Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect. Vertical spacing (`a2`) is given by the height of a hexagon, $\sqrt{3}/2 \times a_1$.

@@

Finally, a tibble is returned containing a unique hexagon ID (`h`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the $2\text{-}D$ space.

>>>>

The coordinate limits of the embedding (`lim1` and `lim2`) are used to compute the aspect ratio between the two axes, which informs vertical spacing. The function then calls `calc_bins_y()`, a helper function that determines the appropriate number of hexagons along y-axis (`b2`) and the width of each hexagon (`a1`) given the specified number of bins along the x-axis (`b1`) and buffer (`q`).

Then, the centroids are computed iteratively. The x-coordinates for centroids in odd-numbered rows are initialized as a sequence spaced by the hexagon width. Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect. Vertical spacing (`vs`) is given by $\sqrt{3}/2 \times a_1$.

@@

Finally, a tibble is returned containing a unique hexagon ID (`h`) along with the corresponding x and y centroid coordinates (`c_x`, `c_y`), which define the layout of the hexagonal grid over the \twoD space.

### 2025-09-18 paper-quollr.tex
Updated vertical spacing

The coordinate limits of the embedding (\texttt{lim1} and \texttt{lim2}) are used to compute the aspect ratio between the two axes, which informs vertical spacing. The function then calls \texttt{calc\_bins\_y()}, a helper function that determines the appropriate number of hexagonal rows (\texttt{bin2}) and the width of each hexagon (\texttt{a1}) given the specified number of bins along the x-axis (\texttt{b1}) and buffer (\texttt{q}).

Then, the centroids are computed iteratively. The x-coordinates for centroids in odd-numbered rows are initialized as a sequence spaced by the hexagon width. Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect. Vertical spacing (\texttt{a2}) is given by the height of a hexagon, \(\sqrt{3}/2 \times a_1\).

@@

Finally, a tibble is returned containing a unique hexagon ID (\texttt{h}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the \(2\text{-}D\) space.

@@

\caption{\label{tab:notation-tb-pdf}Summary of notation for describing new methodology.}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{10cm}>{\raggedright\arraybackslash}p{3cm}}
\toprule
Notation & Description & Variable\\
\midrule
\textbackslash{}\$n, p, k\textbackslash{}\$ & number of observations, variables, embedding dimension, respectively & NA\\
\$\textbackslash{}mathbfit\{X\}, \textbackslash{}mathbfit\{x\}\$ & \$p\$-dimensional data (population, sample) & NA\\
\$\textbackslash{}mathbfit\{y\}\$ & \$k\$-dimensional layout & NA\\
\$P\$ & orthonormal basis, generating a \$d\textbackslash{}text\{-\}dimensional\$ linear projection of \$p\$-dimensional data & NA\\
\$T\$ & true  model & NA\\
\addlinespace
\$g\$ & functional mapping from \textbackslash{}pD\{\} to \textbackslash{}kD\{\}, especially as prescribed by NLDR & NA\\
\$\textbackslash{}mathbfit\{\textbackslash{}theta\}\$ & (Hyper-) parameters for NLDR method & NA\\
\$r\$ & ranges of the embedding components & NA\\
\$C\textasciicircum{}\{(j)\}\$ & \$j\$-dimensional bin centers & NA\\
\$(b\_1, b\_2)\$ & number of bins in each direction & (b1, b2) = bins\\
\addlinespace
\$(a\_1, a\_2)\$ & binwidths, distance between centroids in each direction & a1, a2\\
\$(s\_1, \textbackslash{} s\_2)\$ & starting coordinates of the hexagonal grid & start\_point\\
\$q\$ & buffer to ensure hexgrid covers data, proportion of data range, 0-1 & q\\
\$m\$ & number of non-empty bins & m\\
\$b\$ & number of  hexagons in the grid & b\\
\addlinespace
\$h\$ & hexagonal id & NA\\
\$l\$ & side length & NA\\
\$A\$ & area & NA\\
\$n\_h\$ & bin count & NA\\
\$w\_h\$ & standardised bin counts & NA\\
\addlinespace
NA & threshold to extract high density hexagons & hd\_thresh\\
\bottomrule
\end{tabular}
\end{table}

>>>>

The coordinate limits of the embedding (\texttt{lim1} and \texttt{lim2}) are used to compute the aspect ratio between the two axes, which informs vertical spacing. The function then calls \texttt{calc\_bins\_y()}, a helper function that determines the appropriate number of hexagons along y-axis (\texttt{b2}) and the width of each hexagon (\texttt{a1}) given the specified number of bins along the x-axis (\texttt{b1}) and buffer (\texttt{q}).

Then, the centroids are computed iteratively. The x-coordinates for centroids in odd-numbered rows are initialized as a sequence spaced by the hexagon width. Even-numbered rows are staggered by half this width to achieve a hexagonal tiling effect. Vertical spacing (\texttt{vs}) is given by \(\sqrt{3}/2 \times a_1\).

@@

Finally, a tibble is returned containing a unique hexagon ID (\texttt{h}) along with the corresponding x and y centroid coordinates (\texttt{c\_x}, \texttt{c\_y}), which define the layout of the hexagonal grid over the $2\text{-}D$ space.

@@

\caption{\label{tab:notation-tb-pdf}Summary of notation for describing new methodology.}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{10cm}>{\raggedright\arraybackslash}p{3cm}}
\toprule
Notation & Description & Variable\\
\midrule
\textbackslash{}\$n, p, k\textbackslash{}\$ & number of observations, variables, embedding dimension, respectively & NA\\
\$\textbackslash{}mathbfit\{X\}, \textbackslash{}mathbfit\{x\}\$ & \$p\$-dimensional data (population, sample) & NA\\
\$\textbackslash{}mathbfit\{y\}\$ & \$k\$-dimensional layout & NA\\
\$P\$ & orthonormal basis, generating a \$d\textbackslash{}text\{-\}dimensional\$ linear projection of \$p\$-dimensional data & NA\\
\$T\$ & true  model & NA\\
\addlinespace
\$g\$ & functional mapping from \textbackslash{}pD\{\} to \textbackslash{}kD\{\}, especially as prescribed by NLDR & NA\\
\$\textbackslash{}mathbfit\{\textbackslash{}theta\}\$ & (Hyper-) parameters for NLDR method & NA\\
\$r\$ & ranges of the embedding components & NA\\
\$C\textasciicircum{}\{(j)\}\$ & \$j\$-dimensional bin centers & (c\_x, c\_y)\\
\$(b\_1, b\_2)\$ & number of bins in each direction & (b1, b2) = bins\\
\addlinespace
\$(a\_1, a\_2)\$ & binwidths, distance between centroids in each direction & a1, a2\\
\$(s\_1, \textbackslash{} s\_2)\$ & starting coordinates of the hexagonal grid & start\_point\\
\$q\$ & buffer to ensure hexgrid covers data, proportion of data range, 0-1 & q\\
\$m\$ & number of non-empty bins & m\\
\$b\$ & number of  hexagons in the grid & b\\
\addlinespace
\$h\$ & hexagonal id & NA\\
\$l\$ & side length & NA\\
\$A\$ & area & NA\\
\$n\_h\$ & bin count & NA\\
\$w\_h\$ & standardised bin counts & NA\\
\addlinespace
NA & threshold to extract high density hexagons & hd\_thresh\\
\bottomrule
\end{tabular}
\end{table}
